---
title: "HW04 Part 1: Complete Chapter 3"
author: "Cody Grainger"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 3

- Change "your name" in the YAML header above to your name.

### Section 3.1

Load the tidyverse packages. I've entered the first code chunk for you. **You must enter all subsequent code chunks and run the code.** 

```{r}
library("tidyverse")
```

If you get an error, then you did not install the `tidyverse` package like you were supposed to in the previous assignment. Go back to that assignment and complete that step. (This should not be the case, though.)

### Section 3.2: First steps

Enter your code chunks for Section 3.2 here.

The plot below uses the mpg dataframe to display the negative relationship bewteen engine size (Displacement) and fuel efficiency
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```


Get in the habit of entering a brief description of what each chunk of code does. Enter this above the chunk, like I did in Section 3.1.

#### Section 3.2 Questions

Answer the questions *completely.* Some answers will require you to write code in a code chunk and also to type text before or after the chunk. Other answers will require only code or only text. 

**1:** Run ggplot(data = mpg). What do you see?
A scatterplot showing the relationship between fuel efficiency and engine displacement


**2:** How many rows are in `mpg`? How many columns? **Hint:** Use the `dim()` function. Type `dim()` in the console to learn what `dim()` does.
```{r}
dim(mpg)
```
If I run dim(mpg) I get 234 and 11 for the rows and columns

**3:** What does the `drv` variable describe? Read the help for `?mpg` to find out.
```{r}
?mpg
```
drv refers to the drivetrain setup of a vehicle, such as front wheel drive(f) and rear wheel drive (4)

**4:** Make a scatterplot of `hwy` vs `cyl`.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = hwy, y = cyl))
```


**5:** What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?
The plot has a linear arrangement of data points. 

### Section 3.3: Aesthetic mappings


#### Section 3.3 questions

**1:** What’s gone wrong with this code? Why are the points not blue?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```


**2:** Which variables in `mpg` are categorical? Which variables are continuous? (*Hint:* type `?mpg` to read the documentation for the dataset). How can you see this information when you run mpg?
```{r}
?mpg
```

 The categorial variables are manufacturer model, year, displacement, cylinders, transmissions, drive type, fuel type, and class/type of car. Continouous included hwy and cty mpg.

**3:** Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables? **Note:** You need several code chunks here to answer this question.
```{r}
# With color mapped to class, we see the same chart as above.
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```
```{r}
# When we map size to class, we get a rather hard to read figure as the data points vary so much in size. I also get a message stating "Using size for a discrete variable is not advised".
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
```
```{r}
# When shape is mapped over class, we get a variety of symbols for the differeing classes of cars. suv does not ppear to have a representative symbol.
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```


**4:** What happens if you map the same variable to multiple aesthetics?
```{r}
# When multiple aesthetics are mapped to the same variable such as color and shape, we get a figure with mulitcolored, multishaped data points with the corresponding legend.
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class, shape = class))
```


**5:** What does the stroke aesthetic do? What shapes does it work with? (*Hint:* use `?geom_point`)
```{r}
# Stroke modifies the width of orders on shapes. When set to 5, the shapes become very thick and hard to distinguish from one another. All shapes in my plot were affected by the change in stroke to 5.
?geom_point
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class, color = class, stroke = 5))
```



**6:** What happens if you map an aesthetic to something other than a variable name, like `aes(colour = displ < 5)`?
```{r}
# When color is mapped to displ, we get a gradient of blues as we move from small displ to larger displ.
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = displ, stroke = 1))
```


### Section 3.5: Facets



#### Section 3.5 questions

**1:** What happens if you facet on a continuous variable? **Note:** Write a prediction, then make and run your code chunk.
```{r}
# If I facet on a continuous variable, then the plot will be split up into many more grids than a plot where categorical variables were faceted.

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(hwy ~ displ)

# The figure is unreadable after the facet is applied to hwy and displ as the plot was split up too much with infomeaton on axes overlapping.
```



**2:** What do the empty cells in plot with `facet_grid(drv ~ cyl)` mean? How do they relate to this plot?
```{r}
# The empty sections have no data to be plotted. For example, there are no data for 5 cylinder 4-wheel-drive cars.
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)
```



**3:** What plots does the following code make? What does `.` do?
```{r}
# In the first plot, the "." the plot grid is arranged horizontally and the plot includes hwy, displ, and drv variables.
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

# In the second plot, the "." the plot grid is arranged vertically and the plot includes hwy, displ, and cyl variables.
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```



**4:** Take the first faceted plot below. What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?



**5:** Read `?facet_wrap`. What does `nrow` do? What does `ncol` do? What other options control the layout of the individual panels? Why doesn’t `facet_grid()` have nrow and ncol argument?


**6:** When using `facet_grid()` you should usually put the variable with more unique levels in the columns. Why?

### Section 3.6: Geometric objects



#### Section 3.6.1 Questions

**1:** What geom would you use to draw a line chart? A boxplot? A histogram? An area chart? **Note:** You will have to make some types of these charts in Part 2 of this assignment.


**2:** Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.



**3:** What does `show.legend = FALSE` do? What happens if you remove it?
**Note:** Skip the "Why do you think I used it earlier in the chapter?" question.


**4:** What does the `se` argument to `geom_smooth()` do?


**5:** Will these two graphs look different? Why/why not?


**6:** Recreate the R code necessary to generate the following graphs. **Note:** Enter a separate code chunk for each graph, numbered 1 to 6, starting row-wise at upper left. 

1: upper left.


2: upper right.


3: middle left.


4: middle right.


5: lower left.


6: lower right. *Hint:* Read the help file for `geom_point`. Use `shape = 21`, `stroke = 2`, and `size = 3`.

### Section 3.7: Statistical transformations.

**Skip this section.** We may return to it later.


### Section 3.8 Position adjustments


#### Section 3.8 Questions

**1:** What is the problem with this plot? How could you improve it. *Hint:* Remember the size of the `mpg` data set from earlier in this exercise. Does the number of points seem correct? **Note:** Write the code to improve it. 


**2:** What parameters to `geom_jitter()` control the amount of jittering?


**3:** Compare and contracts `geom_jitter()` with `geom_count()`.


**4:** What is the default position adjustement for `geom_boxplot()`? Create a visualization of the `mpg` dataset that demonstrates it.


### Section 3.9: Coordinate systems

**Note:** Read about and do only the code chunk associated with `coord_flip()`. `coord_flip()` is a great layer to use when you have categorical data with longish names as it makes the text easier to read.

### Section 3.10: The layered grammar of graphics

Just read this section as a contextual review of everything you just learned.


