---
title: "Intro to Basics"
subtitle: "Free Introduction to R"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## 1.1 Comments and calculators

#### Instructions

Here is some sample code. Can you see which lines are actual R code and which are comments?

Add a line of code in the space provided that calculates the sum of 6 and 12, and press the green arrow to run your code chunk.


```{r}
# Calculate 3 + 4
3 + 4

# Calculate 6 + 12

6 + 12

```

## 1.2 Arithmetic with R


#### Instructions

- Type `2 ^ 5` where indicated in the code chunk below to calculate 2 to the power 5.

- Type `28 %% 6` where indicated in the code chunk to calculate 28 modulo 6.

Click the green arrow and have a look at the R output below the code.
Note how the # symbol is used to add comments on the R code but is not included in the outout.


```{r}
# An addition
5 + 5 

# A subtraction
5 - 5 

# A multiplication
3 * 5

 # A division
(5 + 5) / 2 

# Exponentiation
2^5


# Modulo
28%%6


```

## 1.3 Variable assignment

#### Instructions

Complete the code in the chunk such that it assigns the value 42 to the variable `x`, and then view the contents of `x`. You know what to do with the green arrow. Notice that when you ask R to print `x`, the value 42 appears. At least it will if you did this correctly!


```{r}
# Assign the value 42 to x
x=42

# Print out the value of the variable x
x

#
```


## 1.4 Variable assignment (2)

#### Instructions

- Type the following code in the editor: `my_apples <- 5`. This will assign the value 5 to `my_apples`.

- Type `my_apples` below the second comment. This will print out the value of `my_apples`.

Run the code chunk and look at the console: you see that the number 5 is printed. So R now links the variable `my_apples` to the value 5.


```{r}
# Assign the value 5 to the variable my_apples
my_apples <- 5

# Print out the value of the variable my_apples
my_apples

#
```


## 1.5 Variable assignment (3)

#### Instructions

- Assign to `my_oranges` the value 6.

- Add the variables `my_apples` and `my_oranges` and have R simply print the result.

- Assign the result of adding `my_apples` and `my_oranges` to a new variable `my_fruit`.

```{r}
# Assign a value to the variables my_apples and my_oranges
my_apples <- 5
my_oranges <- 10

# Add these two variables together
my_apples + my_oranges 

# Create the variable my_fruit
my_fruit <- my_apples + my_oranges
my_fruit
#
```


## 1.6 Apples and oranges

- Run the code chunk and read the error message. Make sure to understand why this did not work.

- Adjust the code so that R knows you have 6 oranges and thus a fruit basket with 11 pieces of fruit.

```{r}
# Answer

# Assign a value to the variable my_apples
my_apples <- 5 

# Fix the assignment of my_oranges
my_oranges <- "six"

# Create the variable my_fruit and print it out
my_fruit <- my_apples + my_oranges 

my_fruit
```



## 1.7 Basic data types in R

#### Instructions

Assign values to the these variables:

- `my_numeric` variable to 42.

- `my_character` variable to "universe". Note that the quotation marks indicate that "universe" is a character.

- `my_logical` variable to `FALSE.` Note that R is case sensitive!


```{r}
# Assign 42 to my_numeric
my_numeric = 42

# Assign "universe" to my_character
my_character = "universe"

# Assign FALSE to my_logical. Remember: Case matters
my_logical = FALSE
```



## 1.8 What is that data type?

#### Instructions

- Complete the code to show the classes of `my_character` and `my_logical`.

```{r include=TRUE}
# Declare variables of different types
my_numeric <- 42
my_character <- "universe"
my_logical <- FALSE 

# Check class of my_numeric
class(my_numeric)

# Check class of my_character
class(my_character)

# Check class of my_logical
class(my_logical)

#
```

