---
title: "2: Vectors"
subtitle: Introduction to R
output: html_notebook
editor_options:
  chunk_output_type: inline
---

## 2.1 Create a vector

#### Instrutions

Do you still remember what you have learned in the first chapter? Assign the value "Go!" to the variable `vegas`. Remember: R is case sensitive!

```{r}
# Define the variable vegas
vegas = "Go!"
```


## 2.2 Create a vector (2)

#### Instructions

Complete the code such that `boolean_vector` contains the three elements: `TRUE`, `FALSE` and `TRUE` (in that order).

```{r}
numeric_vector <- c(1, 10, 49)
character_vector <- c("a", "b", "c")

# Complete the code for boolean_vector
boolean_vector = c("TRUE", "FALSE", "TRUE")
```


## 2.3 Create a vector (3)

#### Instructions

- Assign the winnings/losses for roulette to the variable `roulette_vector`.

```{r}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday
roulette_vector <- c(140, -50, 20, -120, 240)
```


## 2.4 Naming a vector

#### Instructions

- The code names the elements in `poker_vector` with the days of the week. Add code to do the same thing for `roulette_vector`.

```{r}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday
roulette_vector <- c(-24, -50, 100, -350, 10)

# Assign days as names of poker_vector
names(poker_vector) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

# Assign days as names of roulette_vector
names(roulette_vector) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")


```



## 2.5 Naming a vector (2)

#### Instructions

- A variable `days_vector `that contains the days of the week has already been created for you.

- Use `days_vector` to set the names of `poker_vector` and `roulette_vector`.

```{r}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday
roulette_vector <- c(-24, -50, 100, -350, 10)

# The variable days_vector
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
 
# Assign the names of the day to roulette_vector and poker_vector
roulette_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
poker_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

roulette_vector
poker_vector
```


## 2.6 Calculating total winnings

#### Instructions

- Take the sum of the variables A_vector and B_vector and assign it to `total_vector`.

- Inspect the result by printing out `total_vector`.

```{r}
# Answer

A_vector <- c(1, 2, 3)
B_vector <- c(4, 5, 6)
total_vector <- A_vector + B_vector
# Take the sum of A_vector and B_vector
A_vector + B_vector
  
# Print out total_vector
total_vector



```


## 2.7 Calculating total winnings (2)

#### Instructions

- Assign to the variable `total_daily` how much you won or lost on each day in total (poker and roulette combined).

```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Assign to total_daily how much you won/lost on each day

total_daily <- poker_vector + roulette_vector

total_daily


```


## 2.8 Calculating total winnings (3)

#### Instructions

- Calculate the total amount of money that you have won/lost with roulette and assign to the variable `total_roulette`.

- Now that you have the totals for roulette and poker, you can easily calculate `total_week` (which is the sum of all gains and losses of the week).

- Print out `total_week`.

```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Total winnings with poker
total_poker <- sum(poker_vector)

# Total winnings with roulette
total_roulette <- sum(roulette_vector)

# Total winnings overall
total_week <- total_poker + total_roulette

# Print out total_week
total_week


```


## 2.9 Comparing total winnings

#### Instructions

- Calculate `total_poker` and `total_roulette` as in the previous exercise. Use the `sum()` function twice.

- Check if your total gains in poker are higher than for roulette by using a comparison. Simply print out the result of this comparison. What do you conclude, should you focus on roulette or on poker?

```{r include=TRUE}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Calculate total gains for poker and roulette
total_poker <- sum(poker_vector)
total_roulette <- sum(roulette_vector)


# Check if you realized higher total gains in poker than in roulette 
total_poker
total_roulette
#Higher gains in poker

```

## 2.10 Vector selection: the good times

#### Instructions

- Assign the poker results of Wednesday to the variable `poker_wednesday`.

```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Define a new variable based on a selection


poker_wednesday <- poker_vector[3]

poker_wednesday


```

## 2.11 Vector selection: the good times (2)

#### Instructions

- Assign the poker results of Tuesday, Wednesday and Thursday to the variable `poker_midweek`.

```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Define a new variable based on a selection
poker_midweek = poker_vector[c(2,3,4)]
poker_midweek


```

## 2.12 Vector selection: the good times (3)

#### Instructions

Assign to `roulette_selection_vector` the roulette results from Tuesday up to Friday; make use of `:`.

```{r include=TRUE}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Define a new variable based on a selection
roulette_selection_vector = roulette_vector[c(2:5)]
roulette_selection_vector

```

## 2.12 Vector selection: the good times (4)

#### Instructions

- Select the first three elements in `poker_vector` by using their names: "Monday", "Tuesday" and "Wednesday". Assign the result of the selection to `poker_start`.

- Calculate the average of the values in `poker_start` with the `mean()` function. Simply print out the result so you can inspect it.

```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Select poker results for Monday, Tuesday and Wednesday
poker_vector[c("Monday", "Tuesday", "Wednesday")]
  
# Calculate the average of the elements in poker_start
poker_start <- poker_vector[c("Monday", "Tuesday", "Wednesday")]
mean(poker_start)

```

## 2.13 Selection by comparison - Step 1

#### Instructions

- Check which elements in `poker_vector` are positive (i.e., > 0) and assign this to selection_vector.

- Print out `selection_vector` so you can inspect it. The printout tells you whether you won (`TRUE`) or lost (`FALSE`) any money for each day.

```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Which days did you make money on poker?
selection_vector <- poker_vector[c(140, -50, 20, -120, 240)>0]
  
# Print out selection_vector
selection_vector


```

## 2.14 Selection by comparison - Step 2

#### Instructions

- Print the values contained in `selection_vector`

- Use `selection_vector` in square brackets to assign the amounts that you won on the profitable days to the variable `poker_winning_days`.

```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Which days did you make money on poker?
selection_vector <- poker_vector[c(1:5)>0]

# Print the values contained in selection_vector
selection_vector

# Select from poker_vector these days
poker_winning_days <- selection_vector>0
poker_winning_days

```


## 2.15 Advanced selection

#### Instructions

- Create the variable `selection_vector`, this time to see if you made profit with roulette for different days.

- Assign the amounts that you made on the days that you ended positively for roulette to the variable `roulette_winning_days`. This vector thus contains the positive winnings of `roulette_vector`.

```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Which days did you make money on roulette?
selection_vector <- roulette_vector[c(1:5)>0]
selection_vector
# Select from roulette_vector these days
roulette_winning_days <- selection_vector>0
roulette_winning_days

```

