---
title: "3: Matrices"
subtitle: Introduction to R
output: html_notebook
editor_options:
  chunk_output_type: inline
---

## 3.1 What's a matrix?

#### Instructions

- Construct a matrix with 3 rows containing the numbers 1 up to 9, filled row-wise.

```{r}
# Construct a matrix with 3 rows that contain the numbers 1 up to 9
matrix(1:9, byrow = TRUE, nrow = 3)

#
```

## 3.2 Analyze matrices, you shall

#### Instructions

- Use `c(new_hope, empire_strikes, return_jedi)` to combine the three vectors into one vector. Call this vector `box_office`.

- Construct a matrix with 3 rows, where each row represents a movie. Use the `matrix()` function to do this. The first argument is the `vector box_office`, containing all box office figures. Next, you'll have to specify `nrow = 3` and `byrow = TRUE`. Name the resulting matrix `star_wars_matrix`.

```{r}
# Box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)

# Create box_office
box_office <- c(new_hope, empire_strikes, return_jedi)
# Construct star_wars_matrix
star_wars_matrix <- matrix(box_office, byrow = TRUE, nrow = 3)
star_wars_matrix
#
```


## 3.3 Naming a matrix

#### Instructions

- Use `colnames()` to name the columns of `star_wars_matrix` with the region vector.

- Use `rownames()` to name the rows of `star_wars_matrix` with the titles vector.

- Print out `star_wars_matrix` to see the result of your work.

```{r}
# Box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)

# Construct matrix
star_wars_matrix <- matrix(c(new_hope, empire_strikes, return_jedi), nrow = 3, byrow = TRUE)

# Vectors region and titles, used for naming
region <- c("US", "non-US")
titles <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")

# Name the columns with region
colnames(star_wars_matrix) <- region

# Name the rows with titles
rownames(star_wars_matrix) <- titles

# Print out star_wars_matrix
star_wars_matrix

#
```


## 3.4 Calculating the worldwide box office

#### Instructions

- Calculate the worldwide box office figures for the three movies and put these in the vector named `worldwide_vector`.

- Print the value of `worldwide_vector`.

```{r}
# Construct star_wars_matrix
box_office <- c(460.998, 314.4, 290.475, 247.900, 309.306, 165.8)
star_wars_matrix <- matrix(box_office, nrow = 3, byrow = TRUE,
                           dimnames = list(c("A New Hope", "The Empire Strikes Back", "Return of the Jedi"), 
                                           c("US", "non-US")))

# Calculate worldwide box office figures
worldwide_vector <- rowSums(star_wars_matrix)

# Print out worldwide_vector
worldwide_vector

#
```

## 3.5 Adding a column for the Worldwide box office

#### Instructions

- Add `worldwide_vector` as a new column to the `star_wars_matrix` and assign the result to `all_wars_matrix`. Use the `cbind()` function.

```{r}
# Construct star_wars_matrix
box_office <- c(460.998, 314.4, 290.475, 247.900, 309.306, 165.8)
star_wars_matrix <- matrix(box_office, nrow = 3, byrow = TRUE,
                           dimnames = list(c("A New Hope", 
                                             "The Empire Strikes Back",
                                             "Return of the Jedi"), 
                                           c("US", "non-US")))

# The worldwide box office figures
worldwide_vector <- rowSums(star_wars_matrix)

# Bind the new variable worldwide_vector as a column to star_wars_matrix
all_wars_matrix <- cbind(star_wars_matrix, worldwide_vector)
all_wars_matrix
#
```

## 3.6 Adding a row

#### Instructions

- Use `rbind()` to paste together `star_wars_matrix` and `star_wars_matrix2`, in this order. Assign the resulting matrix to `all_wars_matrix`.

```{r}
# star_wars_matrix and star_wars_matrix2 are available in your workspace
box_office2 <- c(474.5, 552.5, 310.7, 338.7, 380.3, 468.5)
star_wars_matrix2 <- matrix(box_office2, nrow = 3, byrow = TRUE,
                            dimnames = list(c("The Phantom Menace", 
                                              "Attack of the Clones", 
                                              "Revenge of the Sith"), 
                                            c("US", "non-US")))

# Combine both Star Wars trilogies in one matrix
all_wars_matrix <- rbind(star_wars_matrix, star_wars_matrix2)
all_wars_matrix
#
```

## 3.7 The total box office revenue for the entire saga

#### Instructions

- Calculate the total revenue for the US and the non-US region and assign `all_wars_matrix`. You can use the `colSums()` function.
- Print out `total_revenue_vector` to have a look at the results.

```{r}
# all_wars_matrix is available in your Notebook
all_wars_matrix

# Total revenue for US and non-US
total_revenue_vector <- colSums(all_wars_matrix)

# Print out total_revenue_vector
total_revenue_vector

#
```

## 3.8 Selection of matrix elements

#### Instructions

- Select the non-US revenue for all movies (the entire second column of `all_wars_matrix`), store the result as `non_us_all`.

- Use `mean()` on `non_us_all` to calculate the average non-US revenue for all movies. Simply print out the result.

- This time, select the non-US revenue for the first two movies in `all_wars_matrix`. Store the result as `non_us_some`.

- Use `mean()` again to print out the average of the values in `non_us_some`.

```{r}
# all_wars_matrix is available in your workspace
all_wars_matrix

# Select the non-US revenue for all movies
non_us_all <- all_wars_matrix[,2]
non_us_all
# Average non-US revenue
mean(non_us_all)
  
# Select the non-US revenue for first two movies
non_us_some <- all_wars_matrix[1:2,2]
non_us_some
# Average non-US revenue for first two movies
mean(non_us_some)

#
```


## 3.9 A little arithmetic with matrices

#### Instructions

- Divide `all_wars_matrix` by 5, giving you the number of visitors in millions. Assign the resulting matrix to `visitors`.
- Print out `visitors` so you can have a look.

```{r}
# all_wars_matrix is available in your workspace
all_wars_matrix

# Estimate the visitors
visitors <- all_wars_matrix / 5

# Print the estimate to the console
visitors

#
```


## 3.10 A little arithmetic with matrices (2)

#### Instructions

- Divide `all_wars_matrix` by `ticket_prices_matrix` to get the estimated number of US and non-US visitors for the six movies. Assign the result to `visitors`.
- From the `visitors` matrix, select the entire first column, representing the number of visitors in the US. Store this selection as `us_visitors`.
- Calculate the average number of US visitors; print out the result.

```{r}
# all_wars_matrix and ticket_prices_matrix are available for you
all_wars_matrix
ticket_prices <- c(5.0, 5.0, 6.0, 6.0, 7.0, 7.0, 4.0, 4.0, 4.5, 4.5, 4.9, 4.9)
ticket_prices_matrix <- matrix(ticket_prices, ncol = 2, byrow = TRUE,
                               dimnames = list(c("A New Hope",
                                                 "The Empire Strikes Back",
                                                 "Return of the Jedi",
                                                 "The Phantom Menace", 
                                                 "Attack of the Clones", 
                                                 "Revenge of the Sith"),
                                               c("US", "non-US")))


# Estimated number of visitors
visitors <- all_wars_matrix / ticket_prices_matrix
visitors
# US visitors
us_visitors <- visitors[,1]

# Average number of US visitors
mean(us_visitors)

#
```
