---
title: "5: Data frames"
subtitle: Introduction to R
output: html_notebook
editor_options:
  chunk_output_type: inline
---

## 5.1 What is a data frame?

#### Instructions

Run the code below. The data from the built-in example data frame `mtcars` will be printed.

```{r}
# Print out built-in R data frame
mtcars 
```

## 5.2 Quick, have a look at your data set

#### Instructions

Call `head()` on the `mtcars` data set to have a look at the header and the first observations.

```{r}
# Call head() on mtcars
head(mtcars)

#
```

## 5.3 Have a look at the structure

#### Instructions

Investigate the structure of `mtcars`. Make sure that you see the same numbers, variables and data types as mentioned above.

```{r}
# Investigate the structure of mtcars
str(mtcars)

#
```

## 5.4 Creating a data frame

#### Instructions

Use the function `data.frame()` to construct a data frame. Pass the vectors `name`, `type`, `diameter`, `rotation` and `rings` as arguments to `data.frame()`, in this order. Call the resulting data frame `planets_df`.

```{r}
# Definition of vectors
name <- c("Mercury", "Venus", "Earth", "Mars", "Jupiter", "Saturn", "Uranus", "Neptune")
type <- c("Terrestrial planet", "Terrestrial planet", "Terrestrial planet", 
          "Terrestrial planet", "Gas giant", "Gas giant", "Gas giant", "Gas giant")
diameter <- c(0.382, 0.949, 1, 0.532, 11.209, 9.449, 4.007, 3.883)
rotation <- c(58.64, -243.02, 1, 1.03, 0.41, 0.43, -0.72, 0.67)
rings <- c(FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE)

# Create a data frame from the vectors
planets_df <- data.frame(name, type, diameter, rotation, rings)
planets_df
#
```

## 5.5 Creating a data frame (2)

#### Instructions

Use `str()` to investigate the structure of the new `planets_df` variable.

```{r}
# Check the structure of planets_df
str(planets_df)

#
```

## 5.6 Selection of data frame elements

#### Instructions

- From `planets_df`, select the diameter of Mercury: this is the value at the first row and the third column. Simply print out the result.

- From `planets_df`, select all data on Mars (the fourth row). Simply print out the result.

```{r}
# The planets_df data frame from the previous exercise is pre-loaded

# Print out diameter of Mercury (row 1, column 3)
planets_df[1,3]

# Print out data for Mars (entire fourth row)
planets_df[4,]

#
```


## 5.7 Selection of data frame elements (2)

#### Instructions

Select and print out the first 5 values in the `"diameter"` column of `planets_df`.

```{r}
# The planets_df data frame from the previous exercise is available

# Select first 5 values of diameter column
planets_df[1:5,"diameter"]

#
```

## 5.8 Only planets with rings

#### Instructions

- Use the `$` sign to select the rings variable from `planets_df`. Store the vector that results as `rings_vector`.

- Print out `rings_vector` to see if you got it right.

```{r}
# planets_df is pre-loaded in your workspace

# Select the rings variable from planets_df
rings_vector <- planets_df$rings
  
# Print out rings_vector
rings_vector

#
```

## 5.9 Only planets with rings (2)

#### Instructions

The code below selects the `name` column of all planets that have rings. Adapt the code so that instead of only the `name` column, all columns for planets that have rings are selected.

```{r}
# planets_df and rings_vector are pre-loaded in your workspace

# Adapt the code to select all columns for planets with rings
planets_df[rings_vector,]
```

## 5.10 Only planets with rings but shorter

#### Instructions

Use `subset()` on `planets_df` to select planets that have a diameter smaller than Earth. Because the `diameter` variable is a relative measure of the planet's diameter w.r.t that of planet Earth, your condition is `diameter < 1`.

```{r}
# planets_df is pre-loaded in your workspace

# Select planets with diameter < 1
subset(planets_df, subset=diameter<1)

#
```

## 5.11 Sorting

#### Instructions

- Read and run this example. What does the argument `decreasing = TRUE` do?

```{r}
mpg_rank <- order(mtcars$mpg, decreasing = TRUE)
head(mtcars[mpg_rank, ], 8)
```


## 5.12 Sorting your data frame

#### Instructions

- Call `order()` on `planets_df$diameter` (the diameter column of planets_df). Store the result as `positions.`

- Sort `planets_df` with the positions vector as row indexes inside square brackets. Keep all columns. Simply print out the result.

- Call `order()` again on `planets_df$rotation`, with the highest value first. Store the result in `rot_position`.

- Sort `planets_df` as you did above, but use `rot_position` as the row index. Print out the result.

```{r}
# planets_df is pre-loaded in your workspace

# Use order() to create positions
positions<- order(planets_df$diameter)
positions
# Use positions to sort planets_df
positions[order(planets_df$diameter)]

# Use order() to create rotation positions
rot_position <- order(planets_df$rotation)

# Use rot_position to sort planets_df
rot_position[order(planets_df$rotation)]

#
```

