---
title: "6: Lists"
subtitle: Introduction to R
output: html_notebook
editor_options:
  chunk_output_type: inline
---

## 6.1 Lists, why would you need them?

**Nothing to do here.**

## 6.2 Creating a list

#### Instructions

Construct a list, named `my_list`, that contains the variables `my_vector`, `my_matrix` and `my_df` as list components.

```{r}
# Vector with numerics from 1 up to 10
my_vector <- 1:10 

# Matrix with numerics from 1 up to 9
my_matrix <- matrix(1:9, ncol = 3)

# First 10 elements of the built-in data frame mtcars
my_df <- mtcars[1:10,]

# Construct list with these different elements:
my_list <- list(my_vector, my_matrix, my_df)
my_list
#
```


## 6.3 Creating a named list

#### Instructions

- Change the code of the previous exercise (see below) by adding names to the components. Use for `my_vector` the name `vec`, for `my_matrix` the name `mat` and for `my_df` the name `df`.

- Print out my_list so you can inspect the output.

```{r}
# Vector with numerics from 1 up to 10
my_vector <- 1:10 

# Matrix with numerics from 1 up to 9
my_matrix <- matrix(1:9, ncol = 3)

# First 10 elements of the built-in data frame mtcars
my_df <- mtcars[1:10,]

# Adapt list() call to give the components names
my_list <- list(vec=my_vector, mat=my_matrix, df=my_df)

# Print out my_list
my_list

#
```


## 6.4 Creating a named list (2)

#### Instructions

Complete the code below to create `shining_list`; it contains three elements:

- moviename: a character string with the movie title (stored in `mov`)
- actors: a vector with the main actors' names (stored in `act`)
- reviews: a data frame that contains some reviews (stored in `rev`)

Do not forget to name the list components accordingly. The names to use are are `moviename`, `actors` and `reviews`).

```{r}
# The variables mov, act and rev are available

mov<-"The Shining"
act<-c("Jack Nicholson", 
         "Shelley Duvall", 
         "Danny Lloyd", 
         "Scatman Crothers", 
         "Barry Nelson" )

scores <- c(4.5, 4, 5.0)
sources <- c("IMDb1", "IMDb2", "IMDb3")
comments <- c("Best Horror Film I Have Ever Seen",
              "A truly brilliant and scary film from Stanley Kubrick",
              "A masterpiece of psychological horror")
rev <- data.frame(scores, sources, comments)

# Finish the code to build shining_list
shining_list <- list(movies = mov,actors = act,reviews = rev)
shining_list

#
```

## 6.5 Selecting elements from a list

#### Instructions

- Select from `shining_list` the vector representing the actors. Simply print out this vector.

- Select from `shining_list` the second element in the vector representing the actors. Do a printout like before.

```{r}
# shining_list is already pre-loaded in the workspace

# Print out the vector representing the actors
shining_list[[2]]

# Print the second element of the vector representing the actors
shining_list[[2]][2]

#
```


## 6.6 Adding more movie information to the list

#### Instructions

- Complete the code below such that an item named `year` is added to the `shining_list` with the value `1980`. Assign the result to `shining_list_full`.

- Finally, have a look at the structure of `shining_list_full` with the `str()` function.

```{r}
# shining_list, the list containing movie name, actors and reviews, is pre-loaded in the workspace

# We forgot something; add the year to shining_list
year=1980
ext_list <- c(shining_list,year)
shining_list_full <- ext_list
# Have a look at shining_list_full
str(shining_list_full)

#
```

